---
title: "36-315 Lab 07"
author: "Eu Jing Chua"
date: "Due Friday, Oct 11, 2019 (6:30pm)"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---


##  Lab 07:  2-D Continuous Data


***
***

```{r message = F, warning = F}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
```


```{r}
olive <- read_csv("https://raw.githubusercontent.com/mateyneykov/315_code_data/master/data/olive_oil.csv")
```

***
***


#  Problem 0

Using the tidyverse style guide.

```{r}
eujingc_315_theme <-  theme_bw() +
  theme(axis.text = element_text(size = 12),
        plot.title = element_text(size = 20, face = "bold", hjust = 0),
        plot.subtitle = element_text(size = 14, face = "italic", hjust = 0),
        text = element_text(size = 14, face = "bold", color = "darkslategrey"))
```


# Problem 1


a.  
```{r}
olive_sub <- select(olive, area, region, palmitic, palmitoleic, stearic)
```


b.  
```{r}
olive_sub <- olive_sub %>% mutate(
    area = recode(olive_sub$area, "South", "Sardinia", "North"),
    region = recode(olive_sub$region,
                    "North Apulia", "South Apulia", "Calabria",
                    "Sicily", "Umbria", "East Liguria", "West Liguria",
                    "Inland Sardinia", "Coastal Sardinia"))
```



c.  
```{r cache = TRUE}
library(GGally)

ggpairs(olive_sub, aes(color = area))
```

We can see from the boxplots of palmitic and palmitoleic conditioned on area, that the south differs from the other two locations quite significantly in terms of these two variables.  
There also seems to be a positive correlation between palmitic abd palmitoleic.

d.  
```{r}
ggplot(olive_sub,
       aes(x = stearic, y = palmitic)) +
    geom_point(aes(color = region, shape = area)) +
    labs(x = "Stearic", y = "Palmitic",
         color = "Region", shape = "Area",
         title = "Palmitic against Stearic") +
    eujingc_315_theme
```


e.  We can see that the Calabria region and South area tend to be grouped together.



***
***


# Problem 2


a.  
```{r}
ggplot(olive_sub, aes(x = stearic, y = palmitic)) +
    geom_point(aes(color = region, shape = area)) +
    geom_density_2d() +
    labs(x = "Stearic", y = "Palmitic",
         color = "Region", shape = "Area",
         title = "Distribution of Palmitic and Stearic") +
    eujingc_315_theme
```


b.  There are several distinct groups, most notably Calabria and Umbria being tight distinct groups in terms of palmitic and stearic.

c.  

```{r}
ggplot(olive_sub,
       aes(x = stearic, y = palmitic)) +
    geom_bin2d() +
    scale_fill_gradient(low = "yellow", high = "red") +
    geom_point(aes(color = region, shape = area)) +
    guides(col = guide_legend(ncol = 2)) +
    labs(x = "Stearic", y = "Palmitic", fill = "Count",
         color = "Region", shape = "Area",
         title = "Distribution of Palmitic and Stearic") +
    eujingc_315_theme
```




***
***


# Problem 3

```{r message = F, warning = F}
gm <- read_csv("https://raw.githubusercontent.com/mateyneykov/315_code_data/master/data/gapminder.csv")
gm_sub <- gm %>% filter(year > 2000) #data we would actually use.
```


a. 
```{r}
ggplot(gm_sub) +
    geom_point(aes(x = gdpPercap, y = lifeExp, color = continent)) +
    scale_x_log10() +
    labs(x = "log10(GDP (per capita))", y = "Life Expectancy (years)",
         color = "Continent",
         title = "Distribution of log10(GDP) and Life Expectancy") +
    eujingc_315_theme
```


b. Overall, we see a rough positive correlation between the log10 GDP per capita and life expectancy. As the log10 GDP per capita increases, we see an increase in life expectancy. This also holds true conditional on most of the continents, except Africa by looking within each color cluster. We can also see that some clusters tend to have higher life expectancy in general, like Asia and Europe in comparison to Africa.

c. 
```{r}
ggplot(gm_sub, aes(x = log10(gdpPercap), y = lifeExp)) +
    geom_point(aes(color = continent)) +
    geom_smooth(method = "lm", color = "red") +
    geom_smooth(color = "blue") +
    labs(x = "log10(GDP (per capita))", y = "Life Expectancy (years)",
         color = "Continent",
         title = "Distribution of log10(GDP) and Life Expectancy") +
    eujingc_315_theme
```

I prefer the linear regression as it is simpler and fits the data well. The non-parametric regression seems approximately linear too, so I would prefer the simpler model to prevent overfitting.

d. First of all, we can see that both confidence bands do not include the trivial slope of 0, so we can conclude that we reject the hypothesis that there is no linear or non-parametric relationship.  
Secondly, the two confidence bands overlap across all log10 GDP per capita, so we fail to reject the hypothesis that both the linear regression and non-parametric regression is the same.


e. The level parameter sets the level of confidence for which we want the confidence band.  
There is a statistically significant linear relationship for countries in Asia, as the 98% confidence band does not contain the slope of 0.  
We can also conclude that the linear relationships between Asia and Europe fail to be significantly different as their confidence bands overlap across all common regions.

```{r}
ggplot(gm_sub %>% filter(continent %in% c("Asia", "Europe")), 
       aes(x = log10(gdpPercap), y = lifeExp)) + 
  geom_point(aes(color = continent)) +
  geom_smooth(aes(color = continent), method = "lm", level = .98) +
  labs(color = "Continent",
       x = "log10(GDP per Capita)",
       y = "Average Life Expectancy",
       title = "Life Expectancy and GDP per Capita")
```

