---
title: "Maps Demo"
author: "Robin Dunn"
date: "June 19, 2019"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---

```{r, include = FALSE}
library(knitr)

opts_chunk$set(cache=TRUE, autodep=TRUE,
               message=FALSE, warning=FALSE, fig.align="center")
```



#  Adding to Static Maps

We can create static maps (maps of different type such as "terrain", "toner-lite", etc.) with the `ggmap` package:

```{r, warning = FALSE, message = FALSE}
# use the following code to install ggmap
# devtools::install_github("dkahle/ggmap", ref = "tidyup") 
library(tidyverse)
library(ggmap)

pitt <- c(left = -79.944248 - 2, bottom = 40.4415861 - 2,
          right = -79.944248 + 2, top = 40.4415861 + 2)

map_base <- get_stamenmap(pitt, maptype = "toner-lite", zoom = 8)

map_object <- ggmap(map_base, extent = "device")

map_object

map_object2 <- ggmap(map_base) +
  labs(x = "Longitude", y = "Latitude")

map_object2
```


Here, we'll work with airline data from [this GitHub repository](https://github.com/jpatokal/openflights). [Here](https://openflights.org/data.html) you can find out more about the dataset. 

Max Marchi wrote a great summary of using maps in R with ggmap [here](http://www.milanor.net/blog/maps-in-r-plotting-data-points-on-a-map/).  We'll follow this example closely in today's lecture. 

Before we begin, note that this is just one example of how you can add interesting information to maps with `ggmap`.  As long as you have latitude and longitude information, you should be able to add data to maps.  For more interesting examples and for an in-depth description of `ggmap`, see the short paper by David Kahle and Hadley Wickham [here](http://stat405.had.co.nz/ggmap.pdf).

##  Load flight data from GitHub

To load data from GitHub, you should navigate to the raw file, copy the URL, and use `read_csv()`.

```{r, warning = FALSE, message = FALSE}
library(pander)

#  Load and format airports data
airports <- read_csv("https://raw.githubusercontent.com/jpatokal/openflights/master/data/airports.dat",
                     col_names = c("ID", "name", "city", "country", "IATA_FAA", 
                                   "ICAO", "lat", "lon", "altitude", "timezone",
                                   "DST"))

airports %>% as.data.frame() %>% head() %>% pander()

#  Load and format routes data
routes <- read_csv("https://raw.githubusercontent.com/jpatokal/openflights/master/data/routes.dat",
                   col_names = c("airline", "airlineID", "sourceAirport", 
                                 "sourceAirportID", "destinationAirport", 
                                 "destinationAirportID", "codeshare", "stops",
                                 "equipment")) %>%
  filter(sourceAirportID != "\\N",
         destinationAirportID != "\\N") %>%
  mutate(sourceAirportID = as.integer(sourceAirportID),
         destinationAirportID = as.integer(destinationAirportID))

routes %>% as.data.frame() %>% head() %>% pander()
```


##  Manipulating the data to get some custom data.frames

Here, we'll do some data manipulation to obtain the number of arrivals/departures per airport.

```{r, warning = FALSE, message = FALSE}
library(pander)

#  Manipulate the routes data to create two new data.frames
#    one for arrivals, one for departures.
#  Each counts the number of flights arriving/departing from each airport.
departures <- routes %>%
  group_by(sourceAirportID) %>%
  summarise(flights = n())

departures %>% head(3) %>% pander()
```

```{r, warning = FALSE, message = FALSE}
arrivals <- routes %>%
  group_by(destinationAirportID) %>%
  summarise(flights = n())

arrivals %>% head(3) %>% pander()
```

```{r, warning = FALSE, message = FALSE}
#  Merge each of the arrivals/departures data.frames with the airports data.frame above
airportD <- left_join(airports, departures, by = c("ID" = "sourceAirportID"))

airportD %>% head(3) %>% pander()

airportA <- left_join(airports, arrivals, by = c("ID" = "destinationAirportID"))

airportA %>% head(3) %>% pander()
```

```{r, warning = FALSE, message = FALSE}
#  Create data.frame of routes to/from a specific city
my_airport_code <- "PIT"

my_routes <- routes %>% filter(sourceAirport == my_airport_code | 
                               destinationAirport == my_airport_code)

my_routes %>% head(3) %>% pander()
```

`my_routes` now contains information on the source and destination airport IDs. We want to include the latitude and longitude coordinates of each airport.

```{r, warning = FALSE, message = FALSE}
#  Add in relevant information from the airports data.frame
#  Do this in two steps, so that you first pull in the source airport information
#  and then pull in the destination airport information
#  This can be done easily with two calls to left_join()
#  The rest of the code is just formatting
my_airport_data <- my_routes %>%
  left_join(airports, by = c("sourceAirport" = "IATA_FAA")) %>%
  dplyr::select(destinationAirport, lat, lon, timezone) %>%
  rename(source_lat = lat, source_lon = lon, source_timezone = timezone) %>%
  left_join(airports, by = c("destinationAirport" = "IATA_FAA")) %>%
  dplyr::select(source_lat, source_lon, source_timezone, lat, lon, timezone) %>%
  rename(dest_lat = lat, dest_lon = lon, dest_timezone = timezone)

my_airport_data %>% head(3) %>% pander()
```


##  Mapping the data we created

We'll use `ggmap` to create a base map of the data.  Now, we'll add layers of points onto the map.

```{r, warning = FALSE, message = FALSE}
#  Use ggmap to visualize the US flights
library(ggmap)

# first need to get lat/lon bounds of US 
US <- c(left = -125, bottom = 10, right = -67, top = 49)
map <- get_stamenmap(US, zoom = 5, maptype = "toner-lite")

#  Visualize the basic map
ggmap(map, extent = "device")
```

```{r, warning = FALSE, message = FALSE}
#  Add points to the map of departures
#  Each point will be located at the lat/long of the airport
#  The size of the points is proportional to the square root of the number of flights at that airport
mapPoints <- ggmap(map) +
  geom_point(aes(x = lon, y = lat, size = sqrt(flights), color = DST), 
             data = airportD, alpha = .5)

#  Add a custom legend to the plot
mapPointsLegend <- mapPoints +
  scale_size_area(breaks = sqrt(c(1, 5, 10, 50, 100, 500)), 
                  labels = c(1, 5, 10, 50, 100, 500), 
                  name = "departing routes")

mapPointsLegend
```


##  Further Data Manipulation and Facetting

Below, we create a new variable (flight type -- arrival or departure) and use it to create a single data.frame of all arrivals and departures.

We'll then facet on this variable so we can simultaneously visualize both arriving and departing flights.

```{r, warning = FALSE, message = FALSE}
#  Create a data.frame containing both departures and arrivals
#  Do this by creating a "type" variable and then combining the two data.frames
#  We will later be able to use the type variable for facetting
airportD$type <- "departures"
airportA$type <- "arrivals"
airportDA <- bind_rows(airportD, airportA)

#  Create our base plot and add the custom legend
mapPointsDA <- ggmap(map) +
  geom_point(aes(x = lon, y = lat, size = sqrt(flights)), 
             data = airportDA, alpha = .5, color = "blue") +
  scale_size_area(breaks = sqrt(c(1, 5, 10, 50, 100, 500)), 
                  labels = c(1, 5, 10, 50, 100, 500), 
                  name = "routes")

#  Facet on flight type (arrival/departure)
mapPointsDA + facet_grid(. ~ type)
```



***
***
  
  
#  Terrain map

```{r, warning = FALSE, message = FALSE}
library(ggmap) 

# Look up the coordinates for Carnegie Mellon. 
# To do this, I typed "Carnegie Mellon" into Google maps 
# and copied the coordinates from the map's URL.
cmu_coords <- data.frame(lat = 40.4428081, lon = -79.9430128)

map <- get_stamenmap(c(cmu_coords$lon - .005, cmu_coords$lat - .005, 
                       cmu_coords$lon + .005, cmu_coords$lat + .005), 
                     maptype = "terrain", zoom = 16)

ggmap(map) + 
  geom_point(aes(x = lon, y = lat), 
             data = cmu_coords, size = 5, color = "red") +
  labs(x = "Longitude", 
       y = "Latitude",
       title = "Carnegie Mellon Map") +
  theme(plot.title = element_text(hjust = 0.5))
```


***
***
  
  
#  Using Spatial Polygons

  
##  Plotting spatial objects with `geom_polygon()`
  
In `ggplot()`, polygons are just another geometry, making it really easy to add geographic shapes (e.g. corresponding to countries, states, counties, etc.) to maps.

```{r, warning = FALSE, message = FALSE}
#  Note:  The sp package can be really fussy at installation
#  If prompted, do not restart R when installing the package
#  If prompted with "Do you want to install from sources the packages which need compilation?", type 'n'
#install.packages("sp", dependencies = T)
library(sp)
library(ggmap)
us_data <- map_data("state")
county_data <- map_data("county")

us_county_map <- ggplot() + 
  geom_polygon(aes(long, lat, group=group), fill = "blue", size = 4, 
               data = county_data) + 
  geom_polygon(aes(long, lat, group=group), color = 'white', 
               fill = NA, data = us_data) + 
  theme_bw() + theme(axis.text = element_blank(), 
                     axis.title = element_blank())
us_county_map
```



#  Map Projections

##  Use `coord_map()` to specify your map projection

Way back at the beginning of the semester, we learned about using `coord_cartesian()` and `coord_polar()` to specify the coordinates of our plots.  We're (finally) revisiting this idea in the context of geographic mapping.

Using the `coord_map()` function, we can specify what kind of map projection we want to use.

```{r, warning = FALSE, message = FALSE}
us_county_map + coord_map("mercator") 
us_county_map + coord_map("polyconic") 
```


#  Choropleth Maps

Create a choropleth map of the United States that shows the number of musuems in each state.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(sp)
library(ggmap)
museums <- read_csv("https://raw.githubusercontent.com/RobinMDunn/stat315_summer19_data/master/museums.csv")
```

```{r, warning = FALSE, message = FALSE}
us_data <- map_data("state")

museums_by_state <- museums %>% 
  group_by(State) %>%
  summarize(Count = n(),
            log_mean_income = log(mean(Income, na.rm = T)))

us_data <- us_data %>%
  left_join(museums_by_state, by = c("region" = "State"))

ggplot(us_data) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = Count), 
               color = "black") + 
  scale_fill_continuous(low = "green", high = "purple") +
  theme_void() +
  coord_map("polyconic") + 
  labs(title = "Number of Museums Per State")

ggplot(us_data) +
  geom_polygon(aes(x = long, y = lat, group = group, 
                   fill = log_mean_income), color = "black") + 
  scale_fill_continuous(low = "green", high = "purple") +
  theme_void() +
  coord_map("polyconic") + 
  labs(title = "Log Mean Income of Museums Per State")
```


#  Layering (Colored) Points

Use `ggmap` to create a map centered on the continental United States. Add points corresponding to the location of each museum, colored by the type of museum.

```{r, warning = FALSE, message = FALSE}
library(ggmap)
#map <- get_map(location = 'United States', zoom = 4)
US <- c(left = -128, bottom = 24, right = -65, top = 52)
map <- get_stamenmap(US, zoom = 5, maptype = "terrain")

#  Visualize the basic map
ggmap(map) +
  geom_point(aes(x = Longitude, y = Latitude, color = Type), 
             data = museums, alpha = .5) + 
  labs(title = "Museum Locations by Type") + 
  theme_void()
```

#  Filtering

Use the same base map in the previous section to create a map centered on the continental United States. Add points corresponding to natural history museums, sized by their income.

```{r, warning = FALSE, message = FALSE}
nat_hist_museums <- filter(museums, 
                           Type == "NATURAL HISTORY MUSEUM",
                           Income != 0)

ggmap(map) +
  geom_point(aes(x = Longitude, y = Latitude, size = Income), 
             data = nat_hist_museums, alpha = .5) + 
  labs(title = "Natural History Museums by Income") + 
  theme_void()
```

